### 2017-04-13 paper/oopsla17-submission/evaluation-accuracy.tex
more notes from kamalika

\paragraph{Results}
\autoref{fig:accuracy-results} shows the results of our experiment.
%
Localizing the type errors in our benchmarks amounted, on average, to
selecting one of 3 correct locations out of a slice of 10.
%
Our baseline of selecting at random achieves 30\% Top-1
accuracy (58\% Top-3), while \ocaml achieves a Top-1 accuracy of 45\%.
%
Interestingly, one only needs two \emph{random} guesses to outperform
\ocaml, with 47\% accuracy.
%
\sherrloc outperforms both baselines with 56\% Top-1 accuracy (84--86\% Top
3), while \mycroft actually underperforms \ocaml with 38--41\% Top-1
accuracy.
%
Finally, we find that \emph{all} of our classifiers outperform \sherrloc,
ranging from 58--62\% Top-1 accuracy (86--88\% Top-3) for the \linear
classifier to 71--74\% Top-1 accuracy (91\% Top-3) for the \hiddenFH.

Surprisingly, there is little variation in accuracy between classifiers.
With the exception of the \linear model, they all achieve around 70\%
Top-1 accuracy and around 90\% Top-3 accuracy.
%
This suggests that the model they learn is relatively simple.
%
In particular, notice that although the \hiddenT has $50\times$ \emph{fewer}
hidden neurons than the \hiddenFH, it only loses around 2\% accuracy.
% In particular, notice that the \hiddenT only loses around 2\% accuracy
% compared to the \hiddenFH,
%
We also note that our classifiers consistently perform better when
trained on the \FALL programs and tested on the \SPRING programs than
vice versa.
% , they appear to generalize better from the \FALL data.
% FIXME: Why? What is your explanation for this? Is it just sizes of those
% datasets or something qualitative about the program pairs in them?

>>>>

\paragraph{Results}
\autoref{fig:accuracy-results} shows the results of our experiment.
%
Localizing the type errors in our benchmarks amounted, on average, to
selecting one of 3 correct locations out of a slice of 10.
%
Our classifiers consistently outperform the competition, ranging from
58--62\% Top-1 accuracy (86--88\% Top-3) for the \linear classifier to
71--74\% Top-1 accuracy (91\% Top-3) for the \hiddenFH.
%
Our baseline of selecting at random achieves 30\% Top-1
accuracy (58\% Top-3), while \ocaml achieves a Top-1 accuracy of 45\%.
%
Interestingly, one only needs two \emph{random} guesses to outperform
\ocaml, with 47\% accuracy.
%
\sherrloc outperforms both baselines, and comes close to our \linear
classifier, with 56\% Top-1 accuracy (84--86\% Top 3), while \mycroft
actually underperforms \ocaml with 38--41\% Top-1 accuracy.
%
% Finally, we find that \emph{all} of our classifiers outperform \sherrloc,
% ranging from 58--62\% Top-1 accuracy (86--88\% Top-3) for the \linear
% classifier to 71--74\% Top-1 accuracy (91\% Top-3) for the \hiddenFH.

Surprisingly, there is little variation in accuracy between our
classifiers.
%
With the exception of the \linear model, they all achieve around 70\%
Top-1 accuracy and around 90\% Top-3 accuracy.
%
This suggests that the model they learn is relatively simple.
%
In particular, notice that although the \hiddenT has $50\times$ \emph{fewer}
hidden neurons than the \hiddenFH, it only loses around 2\% accuracy.
% In particular, notice that the \hiddenT only loses around 2\% accuracy
% compared to the \hiddenFH,
%
We also note that our classifiers consistently perform better when
trained on the \FALL programs and tested on the \SPRING programs than
vice versa.
% , they appear to generalize better from the \FALL data.
% FIXME: Why? What is your explanation for this? Is it just sizes of those
% datasets or something qualitative about the program pairs in them?

### 2017-04-13 paper/oopsla17-submission/evaluation-utility.tex
more notes from kamalika

\subsection{Feature Utility}
\label{sec:feature-utility}
We have shown that we can train a classifier to effectively localize
type errors, but which of the 276 features that we use are contributing
the most?
%
To answer this question we investigate the performance of
classifiers trained on various subsets of the features.

@@

We tested our hypothesis with the \linear and
%
\hiddenFH\footnote{A layer of 500 neurons is excessive when we have so few
  input features --- we use \hiddenFH for continuity with the
  surrounding sections.}
%
classifiers, cross-validated ($k=10$) over the combined SP14/FA15
dataset. We used a learning rate $\eta=0.001$, $L_2$ regularization rate
$\lambda=0.001$, and mini-batch size of 200. We trained for a single
epoch on feature sets (1) and (2), and for 8 epochs on (3), so that the
total number of training samples would be roughly equal for each feature
set.
%
\lstDeleteShortInline{|} % sigh...
In addition to the accuracy metric from before, we report each
classifier's \emph{recall} --- \ie ``How many true changes can we
remember?'' --- defined as
$$
\frac{|\mathsf{predicted} \cap \mathsf{oracle}|}
     {|\mathsf{oracle}|}
$$
where $\mathsf{predicted}$ is limited to the top 3 predictions of the
classier, and $\mathsf{oracle}$ is limited to the expressions that
changed \emph{and} were part of the type error slice. We make the
latter distinction as (1) changes that are not part of the type error
slice are noise in the data set, and (2) it makes the comparison between
feature sets easier to interpret since $\mathsf{oracle}$ never changes.
% NOTE: keep this at the end of the para or it screws up spacing...
\lstMakeShortInline{|}
%
\input{evaluation-utility-graph}
% \begin{table}[ht]
%   \caption{
%     Impact of Type Error Slice on Accuracy.
%     \ES{TODO: load these numbers from CSV}
%   }\label{tab:type-error-slice}
%   \centering
%   \begin{tabular}{lrcrrrrcrrrr}
%     \toprule
%                        &             & & \multicolumn{4}{c} \linear        & & \multicolumn{4}{c} \hiddenFH      \\
%                                          \cmidrule{4-7}                        \cmidrule{9-12}
%     Feature Set        & \# Features & & Top-1  & Top-2  & Top-3  & Recall & & Top-1  & Top-2  & Top-3  & Recall \\
%     \midrule
%     Local Syntax       & 44          & & 27.7\% & 46.7\% & 58.6\% & 38.2\% & & 32.6\% & 49.5\% & 60.2\% & 39.4\% \\
%     + \InSlice         & 45          & & 46.4\% & 65.1\% & 76.1\% & 48.9\% & & 55.4\% & 71.0\% & 82.1\% & 57.1\% \\
%     Filter by \InSlice & 44          & & 55.9\% & 71.9\% & 82.9\% & 57.5\% & & 57.8\% & 72.7\% & 82.9\% & 57.6\% \\
%     \bottomrule
%   \end{tabular}
% \end{table}

\autoref{fig:slice-utility} shows the results of our experiment.
%
As expected, the baseline performs the worst, with a mere 27.7\% \linear
Top-1 accuracy.
%
Adding \InSlice improves the results substantially with a 46.4\% \linear Top-1
accuracy, demonstrating the importance of a minimal error slice.
%
However, filtering out expressions that are not part of the slice
\emph{further} improves the results to 55.9\% \linear Top-1 accuracy.
%
Interestingly, while the \hiddenFH performs similarly poor with no error
slice features, it recovers nearly all of its accuracy after being given
the error slice features.
%
Top-1 accuracy jumps from 32.4\% to 55.4\% when we add \InSlice, and only
improves by 2\% when we filter out expressions that are not part of the
error slice.

>>>>

\subsection{Feature Utility}
\label{sec:feature-utility}
We have shown that we can train a classifier to effectively localize
type errors, but which of the feature classes from
\autoref{sec:features} are contributing the most to our accuracy?
%
We focus specifically on feature \emph{classes} rather than individual
features as our 276 features are conceptually grouped into a much
smaller number of \emph{categorical} features.
%
For example, the syntactic class of an expression is conceptually a
feature but there are 44 possible values it could take; to encode this
feature for learning we split it into 44 distinct binary features.
%
Analyses that focus on individual features, \eg \textsc{ANOVA},
are difficult to interpret in our setting, as they will tell us the
importance of the binary features but not the higher-level categorical
features.
%
Thus, to answer our question we investigate the performance of
classifiers trained on various subsets of the feature classes.

@@

We tested our hypothesis with the \linear and
%
\hiddenFH\footnote{A layer of 500 neurons is excessive when we have so few
  input features --- we use \hiddenFH for continuity with the
  surrounding sections.}
%
classifiers, cross-validated ($k=10$) over the combined SP14/FA15
dataset. We used a learning rate $\eta=0.001$, $L_2$ regularization rate
$\lambda=0.001$, and mini-batch size of 200. We trained for a single
epoch on feature sets (1) and (2), and for 8 epochs on (3), so that the
total number of training samples would be roughly equal for each feature
set.
%
\lstDeleteShortInline{|} % sigh...
In addition to the accuracy metric from before, we report each
classifier's \emph{recall} --- \ie ``How many true changes can we
remember?'' --- defined as
$$
\frac{|\mathsf{predicted} \cap \mathsf{oracle}|}
     {|\mathsf{oracle}|}
$$
where $\mathsf{predicted}$ is limited to the top 3 predictions of the
classier, and $\mathsf{oracle}$ is the student's fix, limited to changes
that part of the type error slice.
%
We make the latter distinction as (1) changes that are not part of the
type error slice are noise in the data set, and (2) it makes the
comparison between feature sets easier to interpret since
$\mathsf{oracle}$ never changes.
% NOTE: keep this at the end of the para or it screws up spacing...
\lstMakeShortInline{|}
%
\input{evaluation-utility-graph}
% \begin{table}[ht]
%   \caption{
%     Impact of Type Error Slice on Accuracy.
%     \ES{TODO: load these numbers from CSV}
%   }\label{tab:type-error-slice}
%   \centering
%   \begin{tabular}{lrcrrrrcrrrr}
%     \toprule
%                        &             & & \multicolumn{4}{c} \linear        & & \multicolumn{4}{c} \hiddenFH      \\
%                                          \cmidrule{4-7}                        \cmidrule{9-12}
%     Feature Set        & \# Features & & Top-1  & Top-2  & Top-3  & Recall & & Top-1  & Top-2  & Top-3  & Recall \\
%     \midrule
%     Local Syntax       & 44          & & 27.7\% & 46.7\% & 58.6\% & 38.2\% & & 32.6\% & 49.5\% & 60.2\% & 39.4\% \\
%     + \InSlice         & 45          & & 46.4\% & 65.1\% & 76.1\% & 48.9\% & & 55.4\% & 71.0\% & 82.1\% & 57.1\% \\
%     Filter by \InSlice & 44          & & 55.9\% & 71.9\% & 82.9\% & 57.5\% & & 57.8\% & 72.7\% & 82.9\% & 57.6\% \\
%     \bottomrule
%   \end{tabular}
% \end{table}

\paragraph{Results}
\autoref{fig:slice-utility} shows the results of our experiment.
%
As expected, the baseline performs the worst, with a mere 27.7\% \linear
Top-1 accuracy.
%
Adding \InSlice improves the results substantially with a 46.4\% \linear Top-1
accuracy, demonstrating the importance of a minimal error slice.
%
However, filtering out expressions that are not part of the slice
\emph{further} improves the results to 55.9\% \linear Top-1 accuracy.
%
Interestingly, while the \hiddenFH performs similarly poor with no error
slice features, it recovers nearly all of its accuracy after being given
the error slice features.
%
Top-1 accuracy jumps from 32.4\% to 55.4\% when we add \InSlice, and only
improves by 2\% when we filter out expressions that are not part of the
error slice.

### 2017-04-13 paper/oopsla17-submission/overview.tex
misc
more notes from kamalika

\paragraph{\textbf{Challenge (2): What is the ``correct'' expression to blame?}}
As we saw in \autoref{fig:sumList}, type errors admit many different
fixes to different expressions, each resulting in a different program.
%
How is the type checker to judge which expressions are more likely to be
incorrect than others?
%
Prior work has often enlisted expert users to manually judge ill-typed
programs and determine the ``correct'' fix
\citep[\eg][]{Loncaric2016-uk}, but this approach does not scale well to
a dataset large enough to support machine learning.
%
Furthermore, while expert users have intimate knowledge of the type system,
they cannot know in general what novice users \emph{intended} to do.

\paragraph{\textbf{Solution: Use the students' subsequent fixes}}
%
Rather than relying on expert users, we allow our students to do the
work of assigning blame themselves.
%
Software development is an iterative process and our students 
eventually fix their own programs, perhaps after multiple ill-fated
attempts.
%
For each ill-typed program in our dataset, we find the first
\emph{subsequent} program that type checks and declare that to be the
fix.
%
From this pair of an ill-typed program and its fix, we can extract a
\emph{diff} of the abstract syntax trees.
%
For example, suppose our student fixed the |sumList| program in
\autoref{fig:sumList} by replacing |[]| with |0|, the diff would include
only the |[]| expression.
%
Thus we would determine that the |[]| expression (and \emph{not} the
|+| or the recursive call |sumList t|) is to blame.

>>>>

\paragraph{\textbf{Challenge (2): What is the ``correct'' expression to blame?}}
As we saw in \autoref{fig:sumList}, type errors admit many different
fixes to different expressions, each resulting in a different program.
%
How is the type checker to judge which expressions are more likely to be
incorrect than others?
%
Prior work has often enlisted expert users to manually judge ill-typed
programs and determine the ``correct'' fix
\citep[\eg][]{Lerner2007-dt,Loncaric2016-uk}, but this approach does not
scale well to a dataset large enough to support machine learning.
%
Furthermore, while expert users have intimate knowledge of the type system,
they cannot know in general what novice users \emph{intended} to do.

\paragraph{\textbf{Solution: Use the students' subsequent fixes}}
%
Rather than relying on expert users, we follow \citet{Zhang2014-lv} and
allow our students to do the work of assigning blame themselves.
%
Software development is an iterative process and our students 
eventually fix their own programs, perhaps after multiple ill-fated
attempts.
%
For each ill-typed program in our dataset, we find the first
\emph{subsequent} program that type checks and declare that to be the
fix.
%
From this pair of an ill-typed program and its fix, we can extract a
\emph{diff} of the abstract syntax trees.
%
For example, suppose our student fixed the |sumList| program in
\autoref{fig:sumList} by replacing |[]| with |0|, the diff would include
only the |[]| expression.
%
Thus we would determine that the |[]| expression (and \emph{not} the
|+| or the recursive call |sumList t|) is to blame.
